---
title: "initial-analysis"
author: "jonathandash"
date: "2018-08-27"
output: workflowr::wflow_html
---

## Introduction
This analysis is designed to investigate the ALS metrics for the field plots and identify important predictors. Once these are produced we generate useful metrics and produce initial maps.

```{r}
#### Setup project ####
library(here)
library(tidyverse)
library(corrplot)
library(randomForest)
library(caret)
library(party)

#### Read data and merge ####

# Read field data
f.df<-read.csv(here('data', 'field-data.csv'))

# Read Grant's lidar metrics
als.df<-read.csv(here('data', 'allmetrics.csv'))

glimpse(als.df)
glimpse(f.df)

# looks like plot names are not consistent
head(als.df)
head(f.df)

# Check with GP that these als lines are logical... If so proceed bying making a linking identifier.
# GP confirmed on Slack message

# merge datasets
als.df$identifier<-paste('RR', als.df$PlotID, sep='') # Checked with GP and seems reasonable
ref<-merge(f.df, als.df, by.x = 'plot', by.y = 'identifier')

str(ref)

#### review correlations ####

# Make a logic check plot the top height against ALS metrics

plot(pineATH ~ p99, data = ref, xlim = c(5,40), ylim = c(5,40))

# doesn't look great but  they may not know how to measure and calculate top height

plot(pineage~ p99, data = ref, xlim = c(5,40), ylim = c(5,40))
# limited variability within a big range of ages... may need review

preds<-ref[ , 33:110] # define predictors

stand.preds<-ref[ , c(2,3,4,5,6,7,8)] # define predictors related to plantation stand structure. This could 
# allow us to test the hypothesis that ALS metrics are no better than stand structure for prediction.
```


Check correlation matrix for the stand variables
```{r}
# Plot correlations for stand variables
#png(here('out', 'cor-matrix-stand-mets.png'), width = 20, height = 20, 
#         units = "cm", res = 500, pointsize = 12,
#         type = "cairo")
M<-cor(ref[ ,c(9,10,11,12,14,15,  2:8)])
corrplot.mixed(M, upper = "circle", lower = "number",
               tl.pos = "lt", tl.col = "black", tl.offset=1) # too many variables not useful
#dev.off()

png(here('out', 'cor-matrix-stand-mets.png'), width = 20, height = 20, 
         units = "cm", res = 500, pointsize = 12,
         type = "cairo")
M<-cor(ref[ ,c(9,10,11,12,14,15,  2:8)])
corrplot.mixed(M, upper = "circle", lower = "number",
               tl.pos = "lt", tl.col = "black", tl.offset=1) 
dev.off()

```

Plot correlation matrix for height percentiles
```{r}

# Plot correlations for height percentiles
#png(here('out', 'cor-matrix-ht-pct.png'), width = 25, height = 25, 
#    units = "cm", res = 500, pointsize = 12,
#    type = "cairo")
M<-cor(ref[ ,c(9,10,11,12,14,15,  33:44)])
corrplot.mixed(M, upper = "circle", lower = "number",
               tl.pos = "lt", tl.col = "black", tl.offset=1) # too many variables not useful
#dev.off()

png(here('out', 'cor-matrix-ht-pct.png'), width = 25, height = 25, 
    units = "cm", res = 500, pointsize = 12,
    type = "cairo")
M<-cor(ref[ ,c(9,10,11,12,14,15,  33:44)])
corrplot.mixed(M, upper = "circle", lower = "number",
               tl.pos = "lt", tl.col = "black", tl.offset=1) # too many variables not useful
dev.off()

```


```{r}

# Plot correlations for bincentiles
#png(here('out', 'cor-matrix-ht-binc.png'), width = 20, height = 20, 
#    units = "cm", res = 500, pointsize = 12,
#    type = "cairo")
M<-cor(ref[ ,c(9,10,11,12,14,15,  45:50)])
corrplot.mixed(M, upper = "circle", lower = "number",
               tl.pos = "lt", tl.col = "black", tl.offset=1) # too many variables not useful
#dev.off()

# Plot correlations for bincentiles
png(here('out', 'cor-matrix-ht-binc.png'), width = 20, height = 20, 
    units = "cm", res = 500, pointsize = 12,
    type = "cairo")
M<-cor(ref[ ,c(9,10,11,12,14,15,  45:50)])
corrplot.mixed(M, upper = "circle", lower = "number",
               tl.pos = "lt", tl.col = "black", tl.offset=1) # too many variables not useful
dev.off()
```

Correlation matrix for height deciles
```{r}

# Plot correlations for deciles
#png(here('out', 'cor-matrix-ht-deciles.png'), width = 20, height = 20, 
#    units = "cm", res = 500, pointsize = 12,
#    type = "cairo")
M<-cor(ref[ ,c(9,10,11,12,14,15,  51:57)])
corrplot.mixed(M, upper = "circle", lower = "number",
               tl.pos = "lt", tl.col = "black", tl.offset=1) # too many variables not useful
#dev.off()

png(here('out', 'cor-matrix-ht-deciles.png'), width = 20, height = 20, 
    units = "cm", res = 500, pointsize = 12,
    type = "cairo")
M<-cor(ref[ ,c(9,10,11,12,14,15,  51:57)])
corrplot.mixed(M, upper = "circle", lower = "number",
               tl.pos = "lt", tl.col = "black", tl.offset=1) # too many variables not useful
dev.off()
```


Correlation matrix for cov - voxels
```{r}
# Plot correlations for voxels_cov
#png(here('out', 'cor-matrix-vox_cov.png'), width = 20, height = 20, 
#    units = "cm", res = 500, pointsize = 12,
#    type = "cairo")
M<-cor(ref[ ,c(9,10,11,12,14,15,  72:77)])
corrplot.mixed(M, upper = "circle", lower = "number",
               tl.pos = "lt", tl.col = "black", tl.offset=1) # too many variables not useful
#dev.off()

png(here('out', 'cor-matrix-vox_cov.png'), width = 20, height = 20, 
    units = "cm", res = 500, pointsize = 12,
    type = "cairo")
M<-cor(ref[ ,c(9,10,11,12,14,15,  72:77)])
corrplot.mixed(M, upper = "circle", lower = "number",
               tl.pos = "lt", tl.col = "black", tl.offset=1) # too many variables not useful
dev.off()
```

Correlation matric for VCI - SCI voxel metrics
```{r}
# Plot correlations for voxels_VCI_SCI
#png(here('out', 'cor-matrix-vox_VCI_SCI.png'), width = 20, height = 20, 
#    units = "cm", res = 500, pointsize = 12,
#    type = "cairo")
M<-cor(ref[ ,c(9,10,11,12,14,15,  78:86)])
corrplot.mixed(M, upper = "circle", lower = "number",
               tl.pos = "lt", tl.col = "black", tl.offset=1) # too many variables not useful
#dev.off()

png(here('out', 'cor-matrix-vox_VCI_SCI.png'), width = 20, height = 20, 
    units = "cm", res = 500, pointsize = 12,
    type = "cairo")
M<-cor(ref[ ,c(9,10,11,12,14,15,  78:86)])
corrplot.mixed(M, upper = "circle", lower = "number",
               tl.pos = "lt", tl.col = "black", tl.offset=1) # too many variables not useful
dev.off()
```

More correlations for Voxel metrics

```{r}

# Plot correlations for voxels_cc
#png(here('out', 'cor-matrix-vox_cc.png'), width = 25, height = 25, 
#    units = "cm", res = 500, pointsize = 12,
#    type = "cairo")
M<-cor(ref[ ,c(9,10,11,12,14,15,  87:97)])
corrplot.mixed(M, upper = "circle", lower = "number",
               tl.pos = "lt", tl.col = "black", tl.offset=1) # too many variables not useful
#dev.off()

# Plot correlations for voxels_cc
png(here('out', 'cor-matrix-vox_cc.png'), width = 25, height = 25, 
    units = "cm", res = 500, pointsize = 12,
    type = "cairo")
M<-cor(ref[ ,c(9,10,11,12,14,15,  87:97)])
corrplot.mixed(M, upper = "circle", lower = "number",
               tl.pos = "lt", tl.col = "black", tl.offset=1) # too many variables not useful
dev.off()
```

More correlations for Voxel metrics

```{r}

# Plot correlations for voxels_cc
#png(here('out', 'cor-matrix-vox_Pcc-end.png'), width = 25, height = 25, 
#    units = "cm", res = 500, pointsize = 12,
#    type = "cairo")
M<-cor(ref[ ,c(9,10,11,12,14,15,  98:110)])
corrplot.mixed(M, upper = "circle", lower = "number",
               tl.pos = "lt", tl.col = "black", tl.offset=1) # too many variables not useful
#dev.off()

png(here('out', 'cor-matrix-vox_Pcc-end.png'), width = 25, height = 25, 
    units = "cm", res = 500, pointsize = 12,
    type = "cairo")
M<-cor(ref[ ,c(9,10,11,12,14,15,  98:110)])
corrplot.mixed(M, upper = "circle", lower = "number",
               tl.pos = "lt", tl.col = "black", tl.offset=1) # too many variables not useful
dev.off()
```

Looks like a lot of the voxel metrics are correlated with themselves and some conventional metrics like bincentiles.

```{r}
plot(b10~ SCI_2, data = ref) # looks like a lot of the voxel metrics are highly correlated with the lower
```


The response variables suggested by Richard Ewans in an e-mail response are listed below. Suggest these arereviewed by Thomas
 at a later date.
 
1. Average top height of native subcanopy - (Column I; native ATH)
2. Canopy cover of native subcanopy - (Column J; nativeCC)
3. Number of tree fern stems - (Column K; treefernstems)
4. Number of native woody stems (other than tree ferns) - (Column L; nativestems)
5. Number of different native woody species higher than 2m (including tree ferns) - (Column N; nativestemspecies)
6. Total number of different native species - (Column O; nativespecies)

```{r}
# Specify response variables
responses<-ref[ c(9,10,11,12,14,15)]
head(responses)
nrow(responses)
nrow(preds)
```

We used Caret for identifying important variables and feature selection. Do this for or the height of native subcanopy first.

```{r}
#### Rank feature importance ####
# Native ATH first

mod.df<-cbind(responses$nativeATH, preds)
colnames(mod.df)[colnames(mod.df) == 'responses$nativeATH'] <- 'nativeATH'

rf_fit_nat_ATH<-train(nativeATH ~ ., 
                      data = mod.df,
                      method = "ranger")
set.seed(7)
# prepare training scheme
control <- trainControl(method="repeatedcv", number=10, repeats=5)
# train the model
model <- train(nativeATH~., data=mod.df, method="cforest",  trControl=control)
# estimate variable importance
importance <- varImp(model, scale=FALSE)
# summarize importance
print(importance)
# plot importance
plot(importance)

#### Feature Selection ####
set.seed(7)

# define the control using a random forest selection function
control <- rfeControl(functions=rfFuncs, method="cv", number=10)
# run the RFE algorithm
results <- rfe(mod.df[,2:70], mod.df[,1], sizes=c(1:10), rfeControl=control)
# summarize the results
print(results)
# list the chosen features
predictors(results)
# plot the results
plot(results, type=c("g", "o"))
```


Now use the same approach for native canopy cover.

```{r}
#nativeCC

mod.df<-cbind(responses$nativeCC, preds)
colnames(mod.df)[colnames(mod.df) == 'responses$nativeCC'] <- 'nativeCC'

rf_fit_nat_CC<-train(nativeCC ~ ., 
                     data = mod.df,
                     method = "ranger")



# prepare training scheme
control <- trainControl(method="repeatedcv", number=10, repeats=3)
# train the model
model <- train(nativeCC~., data=mod.df, method="cforest",  trControl=control)
# estimate variable importance
importance <- varImp(model, scale=FALSE)
# summarize importance
print(importance)
# plot importance
plot(importance)

#### Feature Selection ####
set.seed(7)

# define the control using a random forest selection function
control <- rfeControl(functions=rfFuncs, method="cv", number=10)
# run the RFE algorithm
results <- rfe(mod.df[,2:70], mod.df[,1], sizes=c(1:8), rfeControl=control)
# summarize the results
print(results)
# list the chosen features
predictors(results)
# plot the results
plot(results, type=c("g", "o"))
```

Use the same approach for Tree fern stems

```{r}
# treefernstems

mod.df<-cbind(responses$treefernstems, preds)
colnames(mod.df)[colnames(mod.df) == 'responses$treefernstems'] <- 'treefernstems'

rf_fit_treeferns<-train(treefernstems ~ ., 
                        data = mod.df,
                        method = "ranger")

# prepare training scheme
control <- trainControl(method="repeatedcv", number=10, repeats=3)
# train the model
model <- train(treefernstems~., data=mod.df, method="cforest",  trControl=control)
# estimate variable importance
importance <- varImp(model, scale=FALSE)
# summarize importance
print(importance)
# plot importance
plot(importance)

#### Feature Selection ####
set.seed(7)

# define the control using a random forest selection function
control <- rfeControl(functions=rfFuncs, method="cv", number=10)
# run the RFE algorithm
results <- rfe(mod.df[,2:70], mod.df[,1], sizes=c(1:8), rfeControl=control)
# summarize the results
print(results)
# list the chosen features
predictors(results)
# plot the results
plot(results, type=c("g", "o"))
```

Number of native stems

```{r}
#native stems

mod.df<-cbind(responses$nativestems, preds)
colnames(mod.df)[colnames(mod.df) == 'responses$nativestems'] <- 'nativestems'

rf_fit_treeferns<-train(nativestems ~ ., 
                        data = mod.df,
                        method = "ranger")

rf_fit_treeferns
# prepare training scheme
control <- trainControl(method="repeatedcv", number=10, repeats=3)
# train the model
model <- train(nativestems~., data=mod.df, method="cforest",  trControl=control)
# estimate variable importance
importance <- varImp(model, scale=FALSE)
# summarize importance
print(importance)
# plot importance
plot(importance)

#### Feature Selection ####
set.seed(7)

# define the control using a random forest selection function
control <- rfeControl(functions=rfFuncs, method="cv", number=10)
# run the RFE algorithm
results <- rfe(mod.df[,2:70], mod.df[,1], sizes=c(1:8), rfeControl=control)
# summarize the results
print(results)
# list the chosen features
predictors(results)
# plot the results
plot(results, type=c("g", "o"))
```

Number of native stem species
```{r}
#nativestemspecies

mod.df<-cbind(responses$nativestemspecies, preds)
colnames(mod.df)[colnames(mod.df) == 'responses$nativestemspecies'] <- 'nativestemspecies'

rf_fit_treeferns<-train(nativestemspecies ~ ., 
                        data = mod.df,
                        method = "ranger")

# prepare training scheme
control <- trainControl(method="repeatedcv", number=10, repeats=3)
# train the model
model <- train(nativestemspecies~., data=mod.df, method="cforest",  trControl=control)
# estimate variable importance
importance <- varImp(model, scale=FALSE)
# summarize importance
print(importance)
# plot importance
plot(importance)

#### Feature Selection ####
set.seed(7)

# define the control using a random forest selection function
control <- rfeControl(functions=rfFuncs, method="cv", number=10)
# run the RFE algorithm
results <- rfe(mod.df[,2:70], mod.df[,1], sizes=c(1:8), rfeControl=control)
# summarize the results
print(results)
# list the chosen features
predictors(results)
# plot the results
plot(results, type=c("g", "o"))
```

Number of native species
```{r}
#nativespecies

mod.df<-cbind(responses$nativespecies, preds)
colnames(mod.df)[colnames(mod.df) == 'responses$nativespecies'] <- 'nativespecies'

rf_fit_treeferns<-train(nativespecies ~ ., 
                        data = mod.df,
                        method = "ranger")

# prepare training scheme
control <- trainControl(method="repeatedcv", number=10, repeats=3)
# train the model
model <- train(nativespecies~., data=mod.df, method="cforest",  trControl=control)
# estimate variable importance
importance <- varImp(model, scale=FALSE)
# summarize importance
print(importance)
# plot importance
plot(importance)

#### Feature Selection ####
set.seed(7)

# define the control using a random forest selection function
control <- rfeControl(functions=rfFuncs, method="cv", number=10)
# run the RFE algorithm
results <- rfe(mod.df[,2:70], mod.df[,1], sizes=c(1:8), rfeControl=control)
# summarize the results
print(results)
# list the chosen features
predictors(results)
# plot the results
plot(results, type=c("g", "o"))
```


